---
title: "Data Mining Final Project"
date: "8/17/2018"
output: html_document
---

# Can load all data frames/model objects built in this analysis by running the below code-chunk
```{r}
#load('DM_Project_vFinal')
```

**-------------------------------------------------------Setup and Data Prep--------------------------------------------------**

```{r}
options(width = 100)
```

```{r, message = FALSE}
library(lubridate)
library(tidyverse)
library(poLCA)
library(MASS)
library(dplyr)
library(rgdal)
library(ggmap)
library(rpart)
```

```{r}
# set your working directory here once
dmprojectpath <- "C:/Users/kshu9_000/Desktop/UChicago/Courses/Quarter 2 - Summer 2018/Data Mining Principles - 31008/Final Project/Datasets/"

df_TSM <- read.csv(paste(dmprojectpath, 'Transportation_Sites_Modified.csv', sep = '/'), stringsAsFactors = F)

# Change the Schools_Serviced Column header name in Bus_Breakdown_and_Delays.csv into OPT_Code before running the below line

df_BBD <- read.csv(paste(dmprojectpath, 'Bus_Breakdown_and_Delays.csv', sep = '/'), stringsAsFactors = F)

df_Routes <- read.csv(paste(dmprojectpath, 'Routes.csv', sep = '/'), stringsAsFactors = F)

df_BBD$Occurred_On <- as.POSIXct(df_BBD$Occurred_On, format="%m/%d/%Y %H:%M")

df_BBD$Created_On <- as.POSIXct(df_BBD$Created_On, format="%m/%d/%Y %H:%M")

df_BBD$Informed_On <- as.POSIXct(df_BBD$Informed_On, format="%m/%d/%Y %H:%M")

df_BBD$Last_Updated_On <- as.POSIXct(df_BBD$Last_Updated_On, format="%m/%d/%Y %H:%M")
```

```{r}
df_TSMBBD <- merge(x = df_TSM, y=df_BBD)
# 174218 obs. of 37 variables

df_clean <- subset(df_TSMBBD, df_TSMBBD$Run_Type != '' & df_TSMBBD$School_Age_or_PreK == 'School-Age')
# 143917 obs. of 37 variables 

df_merge <- as.data.frame(merge(x = df_clean, y = df_Routes))
# 142603 obs. of 49 variables

df_final <- subset(df_merge, df_merge$Boro != '' & df_merge$Boro != 'All Boroughs')
# 135237 obs. of 49 variables
```

**--------------------------------------------------------------LCA------------------------------------------------------------**

## First LCA

### City, Run_Type, Vehicle_TypeDescription, Reason

```{r}
fml1 <- data.frame(df_final[, c("Run_Type", "Reason", "City", "Vehicle_TypeDescription")])

df1 <- as.data.frame(apply(fml1, 2, factor))
```

```{r}
nSample <- 135237
train_size <- round(nSample * (70 / 100))

set.seed(7881)
train_data <- sample(1:nrow(df1), train_size) 
df.train <- df1[train_data, ]
df.holdout <- df1[-train_data,]

i <-  2:5
lca.results.1 <- lapply(i, function(i) {
  f1 <- cbind(City,
              Run_Type,
              Vehicle_TypeDescription,
              Reason) ~ 1
  results <- list()
  results$noOfClasses <- i
  cat("No.ofclasses", i)
  cat("\n\n\n")
  LCA1 <- poLCA(f1, df.train, nclass = i, nrep = 10, tol = .001, verbose = FALSE, graphs = TRUE)
  results$aic <- LCA1$aic
  results$bic <- LCA1$bic
  return(list(results = results, LCA = LCA1))
})
```

```{r}
lca.results.report.1 <- data.frame( do.call(rbind, lapply(lca.results.1, function(x){x$res})) ) 

matplot(lca.results.report.1$noOfClasses,
        cbind(lca.results.report.1$aic, lca.results.report.1$bic),
        type = "b",
        col = c("purple","red"),
        lty = c(1, 1),
        xlab = 'Number of Classes',
        ylab = 'Information Criterion')
legend(x = 'topright', legend = c('AIC','BIC'), lty = c(1, 1), col = c('purple','red'))
```

## Second LCA

### City, Run_Type, Vehicle_TypeDescription, Boro

```{r}
fml2 <- data.frame(df_final[, c("Run_Type", "Boro", "City", "Vehicle_TypeDescription")])

df2 <- as.data.frame(apply(fml2, 2, factor))
```

```{r}
nSample <- 135237 
train_size <- nSample * (70 / 100)

set.seed(7881)
train_data <- sample(1:nrow(df2), train_size) 
df.train <- df2[train_data, ]
df.holdout <- df2[-train_data,]

i <-  2:5
lca.results.2 <- lapply( i, function(i) {
  f2 <- cbind(City,
              Run_Type,
              Vehicle_TypeDescription,
              Boro) ~ 1
  results <- list()
  results$noOfClasses <- i
  cat("No.ofclasses", i)
  cat("\n\n\n")
  LCA2 <- poLCA(f2, df.train, nclass = i, nrep = 10, tol = .001, verbose = FALSE, graphs = TRUE)
  results$aic <- LCA2$aic
  results$bic <- LCA2$bic
  results$probs <- LCA2$probs
  return(list(results = results, LCA = LCA2))
})
```

```{r}
lca.results.report.2 <- data.frame( do.call(rbind, lapply(lca.results.2, function(x){x$res})) ) 

matplot(lca.results.report.2$noOfClasses,
        cbind(lca.results.report.2$aic, lca.results.report.2$bic),
        type = "b",
        col = c("purple","red"),
        lty = c(1, 1),
        xlab = 'Number of Classes',
        ylab = 'Information Criterion')
legend(x = 'topright', legend = c('AIC','BIC'), lty = c(1, 1), col = c('purple','red'))
```

**Holdout validation of chosen solution (3-Class)**
```{r}
f2 <- cbind(City,
            Run_Type,
            Vehicle_TypeDescription,
            Boro) ~ 1
LCA2.3.holdout <- poLCA(f2, df.train, nclass = 3, nrep = 10, tol = .001, verbose = FALSE, graphs = TRUE, 
                        probs.start=lca.results.report.2$probs[2])
```

**When do delays occur for Special Ed AM Runs?**
```{r}
times <- apply((subset(cbind(substr(df_final$Occurred_On, 12, 13),substr(df_final$Occurred_On, 15, 16)), 
                 df_final$Run_Type=='Special Ed AM Run')), 2, as.numeric)
times <- times[,1] + round((times[,2]/60),1)
hist(times)
hist(times[times<7 & times>5.9])
hist(times[times<8 & times>6.9])
```

**How many delays/breakdowns correspond to routes with multiple schools listed?**
```{r}
length(which(grepl("@",unique(df_final$Name))))/length(unique(df_final$Name))
```

**-------------------------------------------------------CART-----------------------------------------------------------------**

**Convert lat and long of school to UTM before building tree model**
```{r}
Trans_Sites <- data.frame(Latitude=df_final$Latitude, Longitude=df_final$Longitude, School.Year=df_final$ï..School_Year)
map <- get_map('New York City', maptype='terrain', source='google')
ggmap(map, base_layer = ggplot(aes(x=Longitude, y=Latitude), data=Trans_Sites)) + geom_point(color="red",alpha=0.05)
cord.dec <- SpatialPoints(cbind(Trans_Sites$Longitude, Trans_Sites$Latitude), proj4string=CRS("+proj=longlat"))
cord.UTM <- spTransform(cord.dec, CRS("+init=epsg:2263"))
plot(cord.UTM, axes=TRUE, xlab="x-coord", ylab="y-coord", main="UTM",pch=1,xlim=c(975000,1000000),ylim=c(100000,350000))
points(df_final$XCoordinates, df_final$YCoordinates, col='green')
```

**Generate Tree 1**
```{r}
unique(df_final$City)
unique(df_final$Garage_City)

df_final.temp <- data.frame(df_final, data.frame(cord.UTM))

df_final.temp$Number_Of_Students_On_The_Bus <- as.numeric(df_final.temp$Number_Of_Students_On_The_Bus)

df_final.temp <- subset(df_final.temp, df_final.temp$Number_Of_Students_On_The_Bus!='NA' & 
                        df_final.temp$Number_Of_Students_On_The_Bus<=50 & df_final.temp$Run_Type=='Special Ed AM Run' 
                        & df_final.temp$Garage_City!='' & df_final.temp$Reason=='Heavy Traffic')
hist(df_final.temp$Number_Of_Students_On_The_Bus,main='Histogram of # students on bus (when delay is reported)')

df_final.temp$Number_Of_Students_On_The_Bus <- (df_final.temp$Number_Of_Students_On_The_Bus==0)*1

df_final.temp$Garage_City[which(df_final.temp$Garage_City=='BROOKLYN')] <- 'Brooklyn'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=='brooklyn')] <- 'Brooklyn'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=='BKLYN')] <- 'Brooklyn'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=='BROOKLYN NY')] <- 'Brooklyn'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="B'klyn")] <- 'Brooklyn'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="BRONX")] <- 'Bronx'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="bronx")] <- 'Bronx'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="BRON")] <- 'Bronx'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="SI")] <- 'Staten Island'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Pelham Manor")] <- 'Westchester'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Pelham")] <- 'Westchester'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="QUEENS")] <- 'Queens'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Jamaica")] <- 'Queens'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Yonkers")] <- 'Westchester'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Astoria")] <- 'Queens'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Whitestone")] <- 'Queens'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Elmont")] <- 'Nassau'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Ozone Park")] <- 'Queens'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Mount Vernon")] <- 'Westchester'
df_final.temp$Garage_City[which(df_final.temp$Garage_City=="Oceanside")] <- 'Nassau'

unique(df_final.temp$City)
unique(df_final.temp$Garage_City)

City.Flag <- (df_final.temp$Garage_City==df_final.temp$Boro)*1

tree.df <- data.frame(City.Flag=factor(City.Flag), Num.S.Flag=factor(df_final.temp$Number_Of_Students_On_The_Bus), 
                      Boro=factor(df_final.temp$Boro), School.Type=factor(df_final.temp$Affiliation))

nSample <- 56855
train_size_tree <- round(nSample * (70 / 100))

set.seed(7881)
train_data_tree <- sample(1:nrow(tree.df), train_size_tree) 

df.train.tree <- tree.df[train_data_tree, ]
df.holdout.tree <- tree.df[-train_data_tree, ]

tree.train <- rpart(df.train.tree,control=rpart.control(cp=0,minsplit=30,xval=10, maxsurrogate=0))
plotcp(tree.train,minline=TRUE,col=4)

cptable <- data.frame(tree.train$cptable)
cp <- cptable[which.min(cptable$xerror),1]

tree.train.pruned <- rpart(df.train.tree,control=rpart.control(cp=cp, minsplit=30, xval=10, maxsurrogate=0))

par(mai=c(0.1,0.1,0.1,0.1))
plot(tree.train.pruned,main="CART: NYC Delays",col=3, compress=TRUE, 
     branch=0.5,uniform=TRUE)
text(tree.train.pruned,cex=0.7,col=4,use.n=TRUE,fancy=TRUE,fwidth=0.2,fheight=0.05,bg=c(5))

table(df.train.tree[,1], predict(tree.train.pruned,type="class"))
round(prop.table(table(df.train.tree[,1], predict(tree.train.pruned,type="class")),1),2)
```

**Generate Tree 2**
```{r}
dist <- sqrt((df_final.temp$XCoordinates-df_final.temp$coords.x1)^2 + (df_final.temp$YCoordinates-df_final.temp$coords.x2)^2)

times <- apply((cbind(substr(df_final.temp$Occurred_On, 12, 13),substr(df_final.temp$Occurred_On, 15, 16))), 2, as.numeric)
times <- times[,1] + round((times[,2]/60),1)

tree.df <- data.frame(City.Flag=factor(City.Flag), 
                      Num.S.Flag=factor(df_final.temp$Number_Of_Students_On_The_Bus), 
                      School.Type=factor(df_final.temp$Affiliation), dist=dist, Boro=factor(df_final.temp$Boro))

nSample <- 56855
train_size_tree <- round(nSample * (70 / 100))

set.seed(7881)
train_data_tree <- sample(1:nrow(tree.df), train_size_tree) 

df.train.tree <- tree.df[train_data_tree, ]
df.holdout.tree <- tree.df[-train_data_tree, ]

tree.train <- rpart(df.train.tree,control=rpart.control(cp=0,minsplit=30,xval=10, maxsurrogate=0))
plotcp(tree.train,minline=TRUE,col=4)

cptable <- data.frame(tree.train$cptable)
cp <- cptable[which.min(cptable$xerror),1]

tree.train.pruned <- rpart(df.train.tree,control=rpart.control(cp=.003, minsplit=30, xval=10, maxsurrogate=0))

par(mai=c(0.1,0.1,0.1,0.1))
plot(tree.train.pruned,main="CART: NYC Bus Delays",col=3, compress=TRUE, 
     branch=0.5,uniform=TRUE)
text(tree.train.pruned,cex=0.7,col=4,use.n=TRUE,fancy=TRUE,fwidth=0.2,fheight=0.02,bg=c(5))

table(df.train.tree[,1], predict(tree.train.pruned,type="class"))
round(prop.table(table(df.train.tree[,1], predict(tree.train.pruned,type="class")),1),2)
```

**Generate Tree 3 - Selected Tree Solution**
```{r}
tree.df <- data.frame(Boro=factor(df_final.temp$Boro), City.Flag=factor(City.Flag), 
                      Num.S.Flag=factor(df_final.temp$Number_Of_Students_On_The_Bus), 
                      School.Type=factor(df_final.temp$Affiliation), dist=dist)

nSample <- 56855
train_size_tree <- round(nSample * (70 / 100))

set.seed(7881)
train_data_tree <- sample(1:nrow(tree.df), train_size_tree) 

df.train.tree <- tree.df[train_data_tree, ]
df.holdout.tree <- tree.df[-train_data_tree, ]

tree.train <- rpart(df.train.tree,control=rpart.control(cp=0,minsplit=30,xval=10, maxsurrogate=0))
plotcp(tree.train,minline=TRUE,col=4)

cptable <- data.frame(tree.train$cptable)
cp <- cptable[which.min(cptable$xerror),1]

tree.train.pruned <- rpart(df.train.tree,control=rpart.control(cp=cp, minsplit=30, xval=10, maxsurrogate=0))

par(mai=c(0.1,0.1,0.1,0.1))
plot(tree.train.pruned,main="CART: NYC Bus Delays",col=3, compress=TRUE, 
     branch=0.5,uniform=TRUE)
text(tree.train.pruned,cex=0.7,col=4,use.n=TRUE,fancy=TRUE,fwidth=0.2,fheight=0.02,bg=c(5))

round(prop.table(table(df.train.tree[,1], predict(tree.train.pruned,type="class")),1),2)
round(prop.table(table(df.holdout.tree[,1], predict(tree.train.pruned,newdata = df.holdout.tree[,-1],type='class')),1),2)
tree.train.pruned$variable.importance
```

```{r}
City.Flag2 <- (df_final.temp$City==df_final.temp$Boro)*1
(table(df_final.temp$Garage_City, df_final.temp$Boro, City.Flag==0, df_final.temp$Number_Of_Students_On_The_Bus==1,
       City.Flag2==1, df_final.temp$Run_Type=='Special Ed AM Run'))
```

**--------------------------------------------Negative Binomial Regression---------------------------------------------------**

```{r}
dfNBR <- df_BBD[,c("Run_Type","Boro","Reason","Occurred_On")]
dfNBR <- subset(dfNBR, dfNBR$Run_Type=="Special Ed AM Run")
dfNBR <- subset(dfNBR, dfNBR$Reason=="Heavy Traffic")
dfNBR <- subset(dfNBR, dfNBR$Boro!='')
dfNBR <- subset(dfNBR, dfNBR$Boro!='All Boroughs')
dfNBR$Occurred_On <- as.POSIXct(dfNBR$Occurred_On, format="%m/%d/%Y")
dfNBR$Date <- format(dfNBR$Occurred_On, "%m/%d/%Y")
dfNBR$weekdays <- weekdays(dfNBR$Occurred_On)

cleaned <- dfNBR[c("Boro", "Date", "weekdays")]
counts <- cleaned %>% group_by(Boro,Date,weekdays) %>% summarise(count_delays=n())
counts$weekdays <- factor(counts$weekdays)
counts$Date <- as.Date(counts$Date, '%m/%d/%Y')

# Verify date/borough combinations with missing counts are holidays
date.range <-seq.Date(as.Date("2015-09-01"), as.Date("2018-06-28"), by="day")
wday <- weekdays(date.range)
date.range.df <- data.frame(date.range, wday)
colnames(date.range.df) <- c("Date", "weekdays")
date.range.df <- subset(date.range.df, date.range.df$weekdays != 'Sunday')
date.range.df <- subset(date.range.df, date.range.df$weekdays != 'Saturday')
Boroughs <- unique(counts$Boro)
date.range.boro.df <- merge(Boroughs, date.range.df, by=NULL)
all.dates <- left_join(date.range.boro.df, counts)
missing.dates <- subset(all.dates, is.na(all.dates$count_delays))
table(missing.dates$Date)

nSample <- 5004
train_size_nbr <- round(nSample * (70 / 100))
set.seed(7881)
train_data_nbr <- sample(1:nrow(counts), train_size_nbr) 
df.train.nbr <- counts[train_data_nbr, ]
df.holdout.nbr <- counts[-train_data_nbr, ]

# Train and Holdout
nb.fit.train <- glm.nb(count_delays ~ Boro + relevel(weekdays,"Monday"), data=df.train.nbr)
summary(nb.fit.train)
RMSE.train <- sqrt(mean((nb.fit.train$y-nb.fit.train$fitted.values)^2))
RMSE.holdout <- sqrt(mean((df.holdout.nbr$count_delays-predict(nb.fit.train, newdata=df.holdout.nbr[,-4],type='response'))^2))
round(cbind(RMSE.train, RMSE.holdout),2)

# Goodness of fit test
pchisq(nb.fit.train$deviance, df=nb.fit.train$df.residual, lower.tail=FALSE) #0.8531

# VS. Poisson
poi.fit.train <- glm(count_delays ~ Boro + relevel(weekdays,"Monday"), data=df.train.nbr, family = poisson(link="log"))
summary(poi.fit.train)

# Predict delay count by day of week, borough
new.data <- merge(unique(counts$Boro), unique(counts$weekdays), by=NULL)
colnames(new.data) <- c("Boro", "weekdays")
predictions <- predict(nb.fit.train, newdata = new.data, type='response')
pred.df <- data.frame(new.data, round(predictions))

Manhattan <- subset(pred.df, pred.df$Boro=='Manhattan')
Queens <- subset(pred.df, pred.df$Boro=='Queens')
Bronx <- subset(pred.df, pred.df$Boro=='Bronx')
Staten.Island <- subset(pred.df, pred.df$Boro=='Staten Island')
Brooklyn <- subset(pred.df, pred.df$Boro=='Brooklyn')

Manhattan <- Manhattan[c(3,1,2,4,5),]
Queens <- Queens[c(3,1,2,4,5),]
Bronx <- Bronx[c(3,1,2,4,5),]
Staten.Island <- Staten.Island[c(3,1,2,4,5),]
Brooklyn <- Brooklyn[c(3,1,2,4,5),]

plot(Manhattan$round.predictions.,ylim=c(0,65), type='l', ylab="Count of Delays", xlab = 'Weekday', xaxt='n', 
     main = "Predicted Count of Delays, By Weekday")
axis(1, at=1:5, labels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
lines(Queens$round.predictions.,type='l',col='red')
lines(Bronx$round.predictions.,type='l',col='blue')
lines(Staten.Island$round.predictions.,type='l',col='green')
lines(Brooklyn$round.predictions.,type='l',col='purple')
legend('topright', legend=c("Manhattan", "Brooklyn", "Bronx", "Queens", "Staten Island"),
       col=c("black", "purple", "blue", "red", "green"), lty=1, cex=0.6)
```
